---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NYCTrash

<!-- badges: start -->
![R-CMD-check](https://github.com/aparikh21/NYCTrash/workflows/R-CMD-check/badge.svg)
<!-- badges: end -->

The goal of NYCTrash is to understand chnages in monthly trash patterns in NYC by borough.

## Installation

You can install the released version of NYCTrash from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("NYCTrash")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("aparikh21/NYCTrash")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(NYCTrash)
library(mosaic)
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


```{r message = FALSE, warning = FALSE, fig.width= 7, fig.width = 7}
NYCTrash
time <- NYCTrash %>%
  group_by(year, borough) %>%
  summarize(refuse_tons_collected = sum(refuse_tons_collected))

ggplot(
  time,
  aes(x = as.numeric(year), y = refuse_tons_collected, color = borough)
) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(1990, 2020, 5)) +
  labs(x = "Year", y = "Tons of Refuse Trash Collected") +
  ggtitle("Changes in Tons of Trash Collected with Time by Borough")
```

Tons of trash collected in NYC have not significantly increased or decreased since the 1990s. 2020 seems to be an unusual year where the tons of trash collected has significantly decreased which could be due to less consumption caused by the pandemic.

Brooklyn is the borough with the most refuse tons of trash collected every year after 1991 while Queens is the borough with the second most tons of refuse trash collected. It is interesting to note that the trash collected in Queens, Brooklyn and Manhattan increase dramatically between 1990 and 1991. It could be worth examining the reason behind this dramatic increase in trash collected. 

The trends during the last 20 years have been similar for all boroughs. All boroughs saw a spike in trash collected in 2003 and all saw declines in trash collected in 2020. The fact that trends have stayed the same between all boroughs could reveal the fact that no borough is significantly improving or becoming worse, hence keeping the ratio of the population in the boroughs roughly similar. 

### b) Is there an association between time of year and the amount of refuse trash collected?

SOLUTION:

```{r message = FALSE, warning = FALSE, fig.width= 7, fig.height= 5}
seasonal_trash <- NYCTrash %>%
  group_by(month, year) %>%
  summarize(refuse = sum(refuse_tons_collected)) %>%
  mutate(month = as.numeric(month)) %>%
  mutate(season = ifelse(month %in% 3:5, "Spring",
                                     ifelse(month %in% 6:8, "Summer", 
                                            ifelse(month %in% 9:11, "Fall", "Winter"))))


ggplot(seasonal_trash,
  aes(x = as.numeric(year), y = refuse, color = season)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", se = FALSE, span = 0.5) + 
  scale_x_continuous(breaks=seq(1990, 2020, 5)) +
  scale_color_manual(breaks = c("Fall", "Spring", "Summer", "Winter"),
                        values=c("tomato3", "limegreen", "deepskyblue3", "black")) +
  labs(x = "Year", y = "Tons of Trash Collected") + 
  ggtitle("Trash Collected in NYC by season since 1990")
```
The winter months of December, January and February have consistently had the lowest amount of trash collected. A possible reason for this could be that people do not go out and consume less during the colder months. On the other hand, the summer months of June, July and August have had the most trash collected suggesting that people consume the most during the warmest months. 

The fall months and spring months have more or less followed similar trends (the decline in the fall line after 2015 is misleading because the line is severely influenced by 2020 fall trash collection.) It would be interesting to investigate the reason behind this drop in tons of trash collected in the fall months of 2020, but it would presumably be because of the pandemic.  

Yearly changes in trash collection seems to affect seasonal consumption evenly because the four lines seem to follow similar trends!
